---
title: "Bing-Je_Wu_HW5"
author: "Bing-Je Wu"
date: "5/6/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Load the data
```{r}
library(curl)
library(RSocrata)
Temp_JSON <- "https://opendata.maryland.gov/resource/pdvh-tf2u.json"
MyJSON_Full <- read.socrata(Temp_JSON)
```

# Step 2: Clean the data
```{r}
MyJSON_org <- data.frame(MyJSON_Full$case_number,MyJSON_Full$barrack,
                         MyJSON_Full$acc_date,MyJSON_Full$acc_time,
                         MyJSON_Full$acc_time_code,MyJSON_Full$day_of_week,
                         MyJSON_Full$road,MyJSON_Full$intersect_road,
                         MyJSON_Full$dist_from_intersect,MyJSON_Full$dist_direction,
                         MyJSON_Full$city_name,MyJSON_Full$county_code,
                         MyJSON_Full$county_name,MyJSON_Full$vehicle_count,
                         MyJSON_Full$prop_dest,MyJSON_Full$injury,
                         MyJSON_Full$collision_with_1,MyJSON_Full$collision_with_2
                         )
```
```{r}
colnames(MyJSON_org)<-c("CASE_NUMBER","BARRACK","ACC_DATE","ACC_TIME","ACC_TIME_CODE",
                         "DAY_OF_WEEK","ROAD","INTERSECT_ROAD","DIST_FROM_INTERSECT",
                         "DIST_DIRECTION","CITY_NAME","COUNTY_CODE","COUNTY_NAME",
                         "VEHICLE_COUNT","PROP_DEST","INJURY","COLLISION_WITH_1",
                         "COLLISION_WITH_2")
```
#analyze the dataset
```{r}
str(MyJSON_org)
```
#normalize the columns
```{r}
MyJSON_org$DAY_OF_WEEK <- gsub('\\ ','',MyJSON_org$DAY_OF_WEEK)
```
#remove NAs
```{r}
colSums(is.na(MyJSON_org))
```
#create a new dataframe without NAs
```{r}
MyJSON_clean <- na.omit(MyJSON_org)
colSums(is.na(MyJSON_clean))
```

# Step 3: Understand the data using SQL (via SQLDF)
## How many accidents happen on SUNDAY
```{r message=FALSE, warning=FALSE}
library(sqldf)
library(gsubfn)
library(proto)
sqldf("select DAY_OF_WEEK, count(*) as Accidents from MyJSON_org 
      where DAY_OF_WEEK='SUNDAY' ")
```
## How many accidents had injuries (might need to remove NAs from the data)
```{r warning=FALSE}
sqldf("select injury, count(*) as Accidents from MyJSON_clean 
      where INJURY = 'YES'")
```
## List the injuries by day
```{r warning=FALSE}
sqldf("select DAY_OF_WEEK,count(injury) as Injuries from MyJSON_org 
      group by DAY_OF_WEEK")
```

# Step 4: Understand the data using tapply
## How many accidents happen on SUNDAY
```{r}
tapply(MyJSON_org$INJURY,MyJSON_org$DAY_OF_WEEK,length)[4]
```
## How many accidents had injuries (might need to remove NAs from the data)
```{r}
tapply(MyJSON_clean$INJURY,MyJSON_clean$INJURY,length)[2]
```
## List the injuries by day
```{r}
tapply(MyJSON_org$INJURY,MyJSON_org$DAY_OF_WEEK, length)
```