---
title: "Google Play Store Apps Academic Project"
author: "Andrea Bradshaw, Bing-Je Wu, Barbara A. Jaehn, Phillip A. Garver"
date: "June 9, 2019"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#  Data Loading & Cleansing
## Load the Google Play Store .CSV File

```{r Prep Libraries & Load Data, message=FALSE, warning=FALSE, paged.print=TRUE}

urlToRead <- "https://raw.githubusercontent.com/bing020815/Syracuse-University/master/IST%20687/Project/data/googleplaystore.csv"
appData<- read.csv(url(urlToRead))
nrow(appData)

```

##  Remove Erroneous Row Import and any Duplicates
```{r Delete bad row, message=FALSE, warning=FALSE, paged.print=FALSE}
appData <- appData[-which(appData$Installs=="Free"),]
paste0("Before: ", nrow(appData))
appData <- unique(appData)
paste0("After: ", nrow(appData))

```

##  Rename All Fields to Use Underscores Between Words Instead of Dots (.)
```{r Rename fields, message=FALSE, warning=FALSE, paged.print=FALSE}
colnames(appData) <- c("App", "Category", "Rating", "Reviews", "Size", "Installs", 
                       "Type","Price", "Content_Rating", "Genre", "Last_Updated", 
                       "Current_Ver","Android_Ver")

#Renumber rows
row.names(appData) <- NULL
paste0("Count After: ", nrow(appData))

```
  
##  Identify and remove all N/A and NAN fields
```{r Remove N/A fields, message=FALSE, warning=FALSE, paged.print=FALSE}
colSums(is.na(appData))

appData <- na.omit(appData)
appData <- appData[-which(appData$Android_Ver =="NaN"),]

#Remove the unused Factor Levels in all columns
appData <- droplevels(appData)

#Confirm N/As have been removed
paste0("Count After: ", nrow(appData))
colSums(is.na(appData))
```


## Convert Size Field to Kilobytes
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Split dataset into three group SizeM SizeK NotSize
SizeM <- appData[grepl(pattern = "M",appData$Size),]
NotSize <- appData[!grepl(pattern = "M",appData$Size),]
SizeK <- NotSize[grepl(pattern = "k",NotSize$Size),]
NotSize <-NotSize[!grepl(pattern = "k",NotSize$Size),]
# verify the number of records
ifelse (nrow(appData) == (nrow(SizeM) + nrow(NotSize) + nrow(SizeK)), 
        "Record Counts Match", "Record Count Discrepancy")

# Converting everything into KB
SizeM$Size <- gsub("\\M","",SizeM$Size)
SizeM$Size <- as.numeric(SizeM$Size)
SizeM$Size <- SizeM$Size*1024
SizeK$Size <- gsub("\\k","",SizeK$Size)
SizeK$Size <- as.numeric(SizeK$Size)
appData <- data.frame(rbind(SizeM,SizeK,NotSize))
#Remove + from Size field data

paste0("Count After: ", nrow(appData))
```


## Remove + in Installs and change into a number (so the buckets to sort numerically)
```{r Convert Installs columns for sorting, message=FALSE, warning=FALSE, paged.print=FALSE}
options("scipen"=100, "digits"=4)
appData$Installs <- gsub("\\+","",appData$Installs)
appData$Installs <- gsub("\\,", "", appData$Installs)
appData$Installs <- as.numeric(appData$Installs)
#appData <- within(appData, {Installs <- as.numeric(as.character(Installs))} )

paste0("Count After: ", nrow(appData))
```

## Convert Reviews to a number
```{r Convert Reviews to a number, message=FALSE, warning=FALSE, paged.print=FALSE}
appData$Reviews <- as.numeric(as.character(appData$Reviews))
#appData <- within(appData, {Reviews <- as.numeric(as.character(Reviews))})

paste0("Count After: ", nrow(appData))
```

## Remove Dollar Sign from Price Field and change to numeric data for Price field 
```{r Clean up Price Data, message=FALSE, warning=FALSE, paged.print=FALSE}
#Put Zeroes into double format and remove $  
appData$Price <- gsub("\\$", "", appData$Price)
appData$Price <- as.numeric(appData$Price)

paste0("Count After: ", nrow(appData))
```

## Identify duplicate entries for Apps and take the Max Reviews and all remaining data as all other columns are the same for these enries.
```{r message=FALSE, warning=FALSE}
library(dplyr)
appData <- appData %>%
  
            group_by(App, Rating) %>%
  
            arrange(desc(Reviews)) %>%
  
            slice(1) %>%
  
            ungroup()

#appData <- appData %>% distinct(App,Rating, .keep_all = TRUE)

##reference:  https://stackoverflow.com/questions/24237399/how-to-select-the-rows-with-maximum-values-in-each-group-with-dplyr
##reference:  https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/
```

## Remove Last_Updated, Current_Ver and Android_Ver Fields
```{r Eliminate Last_Updated & Current_Ver Fields, message=FALSE, warning=FALSE, paged.print=TRUE}
appData <- appData[, -11:-13]
paste0("Count After: ", nrow(appData))

```

## Break out Sub-Genre's Where Multiple Genre Entries Exist in the Same Field, then set NAs to None
```{r Break out Sub-Genres, message=FALSE, warning=FALSE, paged.print=TRUE}
library(tidyr)
appData <- separate(appData, Genre, into = c("Genre", "SubGenre"), sep = ";")

library(sqldf)
#Show Number of apps with a SubGenre:
sqldf('select count(*) from appData where SubGenre is NOT NULL')

#Replace SubGenre NA values with None
appData$SubGenre[is.na(appData$SubGenre)] = "None"

paste0("Count After: ", nrow(appData))
str(appData)
```

## Preliminary Prep of RandomForsest Data Set, and Adding SizeCategories
```{r message=FALSE, warning=FALSE}
#Remove App column from target dataset for randomForest analysis
rfappData <- data.frame(appData[,-1])

#Create SizeCategory column
rfappData$SizeCategory <- NULL

#Switch to SizeCategories to include Varies by device
rfappData$SizeCategory[rfappData$Size == "Varies with device"] = "Varies with device"
rfappData$SizeCategory[as.numeric(is.na(rfappData$Size))] = "Varies with Device"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 0 & 
                         as.numeric(rfappData$Size) <= 1024] = "0-1MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 1024 & 
                         as.numeric(rfappData$Size) <= 5120] = "1MB-5MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 5120 & 
                         as.numeric(rfappData$Size) <= 10240] = "5MB-10MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 10240 & 
                         as.numeric(rfappData$Size) <= 25600] = "10MB-25MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 25600 & 
                         as.numeric(rfappData$Size) <= 51200] = "25MB-50MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 51200 & 
                         as.numeric(rfappData$Size) <= 76800] = "50MB-75MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 76800 & 
                         as.numeric(rfappData$Size) <= 102400] = "75MB-100MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 102400 & 
                         as.numeric(rfappData$Size) <= 128000] = "100MB-125MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 128000 & 
                         as.numeric(rfappData$Size) <= 153600] = "125MB-150MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 153600 & 
                         as.numeric(rfappData$Size) <= 179000] = "150MB-175MB"
rfappData$SizeCategory[as.numeric(rfappData$Size) > 179000 & 
                         as.numeric(rfappData$Size) <= 204800] = "175MB-200MB"

str(rfappData)
```





# Display initial descriptive statistics & visualizations of all elements.

## Summarize our data

```{r Summarize Data, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(appData[,2:9])
```

## Show Summary of App Reviews
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
summary(appData$Reviews)
```

## Preliminary Data Visualization

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(reshape2)
ggplot(data = melt(appData), mapping = aes(x = value)) + 
  geom_histogram(bins=20, color = "black", fill = "green") + 
  facet_wrap(~variable, scales = 'free_x') + theme(axis.text.x = element_text(angle=45, hjust=1))
```

## Average Rating of all Apps
```{r}
mr <- mean(appData$Rating)
mr

mrPlot <- ggplot(appData, aes(x=Rating)) + 
  geom_bar(fill = "blue")  + 
  ggtitle("Rating Distribution of Apps") + 
  xlab("Ratings") + ylab("App Count") +
  theme(plot.title = element_text(hjust = 0.5))
mrPlot
```

## Average Rating by Category
```{r}
mr1 <- data.frame(tapply(appData$Rating, appData$Category, mean))
mr1 <- cbind(rownames(mr1), data.frame(mr1, row.names = NULL))
colnames(mr1) <- c("Category", "AverageRating")


mrPlot1 <- ggplot(mr1, aes(x=reorder(Category,AverageRating), y = AverageRating)) + 
  geom_bar(stat="identity", fill = "blue") + 
  ggtitle("Average Rating by Category") + xlab("Category") + 
  coord_flip() +
  theme(plot.title = element_text(hjust = 0.5))  

maxCatRating <- mr1[which.max(mr1$AverageRating),]

paste0("Category with Maximum Average Rating: ", maxCatRating$Category, "   Average Rating: ", maxCatRating$AverageRating)

mrPlot1


```

## Average Rating by Genre
```{r}
mr2 <- data.frame(tapply(appData$Rating, appData$Genre, mean))
mr2 <- cbind(rownames(mr2), data.frame(mr2, row.names = NULL))
colnames(mr2) <- c("Genre", "AverageRating")


mrPlot2 <- ggplot(mr2, aes(x=reorder(Genre,AverageRating), y = AverageRating)) + 
  geom_bar(stat="identity", fill = "blue") + 
  ggtitle("Average Rating by Genre") + 
  xlab("Genre") + 
  coord_flip() +
  theme(plot.title = element_text(hjust = 0.5)) 

maxGenreRating <- mr2[which.max(mr2$AverageRating),]

paste0("Genre with Maximum Average Rating: ", maxGenreRating$Genre, "   Average Rating: ", maxGenreRating$AverageRating)

mrPlot2
```

## Genre by Installs
```{r message=FALSE, warning=FALSE}
ggplot(appData, aes(x=Genre, y = Installs)) + 
  geom_bar(stat="identity", color = "red", fill = "red") + 
  ggtitle("Genre by Installs") + 
  theme(axis.text.x = element_text(angle=90, hjust=1), plot.title = element_text(hjust = 0.5))
```
## Category by Installs
```{r message=FALSE, warning=FALSE}
ggplot(appData, aes(x=Category, y = Installs)) + 
  geom_bar(stat="identity", color = "red", fill = "red") + 
  ggtitle("Category by Installs") + 
  theme(axis.text.x = element_text(angle=90, hjust=1), plot.title = element_text(hjust = 0.5))
```

## Installs by Ratings
```{r message=FALSE, warning=FALSE}
ggplot(appData, aes(x=Rating, y = Installs)) + 
  geom_bar(stat="identity", color = "red", fill = "red") + 
  ggtitle("Installs by Rating") + 
  theme(plot.title = element_text(hjust = 0.5))
```

## Installs by Content_Rating
```{r message=FALSE, warning=FALSE}
ggplot(appData, aes(x=Content_Rating, y = Installs)) + 
  geom_bar(stat="identity", color = "red", fill = "red") + 
  ggtitle("Installs by Content Rating") + 
  theme(axis.text.x = element_text(angle=90, hjust=1), plot.title = element_text(hjust = 0.5))
```

## SubGenre impact on Installs
```{r message=FALSE, warning=FALSE}
ggplot(appData, aes(x=SubGenre, y = Installs)) + 
  geom_bar(stat="identity", color = "red", fill = "red") + 
  ggtitle("Installs by SubGenre") + 
  theme(axis.text.x = element_text(angle=90, hjust=1), plot.title = element_text(hjust = 0.5))
```

## Size Category by Installs with Rating
```{r Size Category by Installs with Rating}
ggplot(rfappData, aes(x=SizeCategory, y = Installs)) + 
  geom_point(aes(color = Rating, size = Rating)) + 
  scale_color_gradient2(midpoint=3, low="yellow", mid="green", high="blue") + 
  theme(axis.text.x = element_text(angle=90, hjust=1), plot.title = element_text(hjust = 0.5)) + 
  ggtitle("SizeCategory by Installs with Rating")
```

## Price by Content_Rating
```{r message=FALSE, warning=FALSE}
ggplot(appData, aes(x=Content_Rating, y = Price, size = Price, color = Price)) + 
  geom_point(stat="identity") + 
  scale_color_gradient(low="red", high="blue") + 
  theme(axis.text.x = element_text(angle=90, hjust=1), plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Price by Content_Rating") 
```

## Top 10 Apps' Earning Categories with Price (Excluding in game Purchases)
```{r Top 10 Apps Earning Categories, message=FALSE, warning=FALSE, paged.print=FALSE}
# Calculate the minimum amount earned by game purchases alone:
appData$MinEarned <- appData$Price * appData$Installs

# Identify Top 10 Earners from game purchases/minimum installs alone
topEarners <- appData[order(appData$MinEarned, decreasing=TRUE), ]
head(topEarners, 10)

#Visualization
ggplot(head(topEarners, 10), aes(x=Category, fill=factor(Rating))) + 
  geom_bar() + 
  ggtitle("Top 10 Earning Apps' Categories") + 
  theme(axis.text.x = element_text(angle=90, hjust=1), plot.title = element_text(hjust = 0.5))

```


## Correlation checks:
```{r message=FALSE, warning=FALSE}
#install.packages("devtools")
#install.packages("ggpubr")
library(devtools)
library(ggpubr)

# Correlation between Ratings and Installs
riCorr <- data.frame(appData$Rating, appData$Installs)
riCorrTest <- cor.test(riCorr$appData.Rating, riCorr$appData.Installs, method="pearson")
riCorrTest
```
Installs and Rating are not correlative

## Average Size of an application (where size does not vary by device):
```{r message=FALSE, warning=FALSE}
ap2 <- appData
ap2$Size <- as.numeric(ap2$Size)
ap2 <- na.omit(ap2)
avgSize <- mean(ap2$Size)
paste0("Average Size of an application: ", avgSize, " KB ")
```





# Modeling for Random Forest and Support Vector Machine

## Prep for randomForest and Run
```{r Prep for randomForest and Run, message=FALSE, warning=FALSE, paged.print=FALSE}
#Remove Size
rfappData <- rfappData[,-4]

# Convert varibles into factor for rfappData
cols <- c("Category","Type","Genre","SubGenre","SizeCategory")
for (i in cols){
  rfappData[,i]<-as.factor(rfappData[,i])
}

# Convert varibles into numeric for rfappData
cols <- c("Rating","Reviews","Installs","Price")
for (i in cols){
  rfappData[,i]<-as.numeric(rfappData[,i])
}

str(rfappData)

#Create Random Forest model splitting categories into >= 100000 installs and below
library(randomForest)
model2 <- randomForest(rfappData[,-4], as.factor(rfappData[,4] >= 100000))
model2

importance(model2)
varImpPlot(model2)
```

## We see the highest ranking independent variables are Reviews, Genre, SizeCategory and then Rating
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Set varibles to reflect lessons learned above
num_exmps = nrow(rfappData)
L = replace(integer(num_exmps), rfappData[,4]>=100000, 1)
M <- rfappData[,-4]

# Use Cross validation to build model
train_idx <- sample(c(1:num_exmps), size = num_exmps * 0.7, replace = FALSE)
model2 <- randomForest(M[train_idx,],as.factor(L[train_idx]))
model2

# Generate propsoed answers using Cross validation
pred <- predict(model2, M[-train_idx,],type="prob")

# Plot ROC metric
library(pROC)
plot(roc(L[-train_idx], as.numeric(pred[,1])))
# ROC info https://en.wikipedia.org/wiki/Receiver_operating_characteristic
```

## Run K-Fold to for cross validation to try to prevent over fitting
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#install.packages("caret")
#install.packages("e1071")
library(caret)
library(e1071)

# Set up cross-validation for k=10 folds
train_Control <- trainControl(method="cv", number=10)
# Train the model with K-Fold cross validation training set
model <- train(M[train_idx,],as.factor(L[train_idx]), trControl=train_Control, method="rf")
print(model)


# Generate propsoed answers using Cross validation
pred <- predict(model, M[-train_idx,],type="prob")

# Plot ROC metric
library(pROC)
plot(roc(L[-train_idx], as.numeric(pred[,1])))
# ROC info https://en.wikipedia.org/wiki/Receiver_operating_characteristic
# Caret Package Reference: https://topepo.github.io/caret/available-models.html
```

## Support Vector Machine
```{r Support Vector Machine, message=FALSE, warning=FALSE, paged.print=FALSE}
library(kernlab)
library(pROC)

svmappData <- data.frame(rfappData)
svmappData$Installs <- ifelse(test = svmappData$Installs>100000, yes = 1, no = 0)

str(svmappData)
names(svmappData)

# Feature analysis from random forest:
# rating > size > genre > review > androvid > Category > price > type > content > rating >"subgenre"

svmappData <- svmappData[,-8] #remove subgenre

## create function
cPercent <- function(predicted, actual){
 confMatrix<- table(predicted, actual, dnn=c("Prediction","Actual"))
 Result <- (confMatrix[1,1]+confMatrix[2,2])/sum(colSums(confMatrix))*100
 print(confMatrix)
 return(sprintf("Correct Percentage: %1.2f%% ", Result))
}  

# create a randomized index
randIndex <- sample(1:nrow(svmappData))

# Calculate the cut point and divide the data set into training set & test set:
cutPoint2_3 <- floor(2*nrow(svmappData)/3)
cutPoint2_3

# generate test set and training data sets:
trainData <- svmappData[randIndex[1:cutPoint2_3],]
testData <- svmappData[randIndex[(cutPoint2_3+1):nrow(svmappData)],]

2737+5474

# Generate a model based on the training data set:
# model 1 --- Radial Basis kernel "Gaussian"
svmOutput <- ksvm(Installs~., data = trainData, kernel = "rbfdot", kpar="automatic", 
                  C=5, cross=3, prob.model=TRUE)
svmOutput

predSVM <- round(predict(svmOutput,testData))
cPercent(predSVM, testData$Installs)
plot(roc(predSVM, testData$Installs))

# model 2 --- Linear kernel
svmOutput2 <- ksvm(Installs~., data = trainData, kernel = "vanilladot", kpar="automatic", 
                   C=5, cross=3, prob.model=TRUE)
svmOutput2

predSVM2 <- ifelse(round(predict(svmOutput2,testData))>=1, 1,0)
cPercent(predSVM2, testData$Installs)
plot(roc(predSVM2, testData$Installs))

# model 3 --- Polynomial kernel
svmOutput3 <- ksvm(Installs~., data = trainData, kernel = "polydot", kpar="automatic", 
                   C=5, cross=3, prob.model=TRUE)
svmOutput3

predSVM3 <- ifelse(round(predict(svmOutput3,testData))>=1, 1,0)
cPercent(predSVM3, testData$Installs)
plot(roc(predSVM3, testData$Installs))
```

Random Forest Model is the best model according to ROC plot and Accuracy Rate
