---
title: "Bing-Je_Wu_HW6"
author: "Bing-Je Wu"
date: "5/12/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
fig_caption: yes
xelatex: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Step 1: Load the data
```{r}
Mydf <- airquality
```

# Step 2: Clean the data
```{r}
colSums(is.na(Mydf))
str(Mydf)
```
There are two options to deal with missing values  
option 1. Using na.omit to remove records with missing values  
option 2. Using na.rm on the each function to exclude missing value from analysis  
In order to get the most valuable analysis, option 2 is the best solution.   

# Step 3 -1: Understand the data distribution
## Create the following visualizations using ggplot:
• Histograms for each of the variables

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
```
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Ozone)) + 
  geom_histogram(binwidth =20,color="black",fill="white", na.rm = TRUE) +
  ggtitle("Ozone Histogram") + theme(plot.title = element_text(hjust = 0.5))
```
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Solar.R)) + 
  geom_histogram(binwidth =50,color="black",fill="white", na.rm = TRUE) +
  ggtitle("Solar.R Histogram") + theme(plot.title = element_text(hjust = 0.5))
```
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Wind)) + 
  geom_histogram(binwidth =1, color="black",fill="white", na.rm = TRUE) +
  ggtitle("Wind Histogram") + theme(plot.title = element_text(hjust = 0.5))

```
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Temp)) + 
  geom_histogram(binwidth=5, color="black",fill="white", na.rm = TRUE) +
  ggtitle("Temp Histogram") + theme(plot.title = element_text(hjust = 0.5))
```
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Month)) + 
  geom_histogram(bins=5, color="black",fill="white", na.rm = TRUE) +
  ggtitle("Month Histogram") + theme(plot.title = element_text(hjust = 0.5))
```
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Day)) + 
  geom_histogram(bins=31,color="black",fill="white", na.rm = TRUE) +
  ggtitle("Day Histogram") + theme(plot.title = element_text(hjust = 0.5))
```

• Boxplot for Ozone
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=factor(0), y=Ozone)) +geom_boxplot(na.rm = TRUE) +
  ggtitle("Ozone Boxplot") + theme(plot.title = element_text(hjust = 0.5))
```

• Boxplot for wind values (round the wind to get a good number of “buckets”)
```{r fig.width = 5, fig.height = 4}
Mydf$Wind_new <- round(Mydf$Wind)
ggplot(Mydf, aes(x=Wind_new, y=Ozone, group=Wind_new)) + geom_boxplot(na.rm = TRUE) + 
  labs(subtitle = "Wind(rounded) vs Ozone Boxplot") + xlab("Wind")
```

# Step 3-2: Explore how the data changes over time
## Combine Month and Day as Date
```{r fig.width = 5, fig.height = 4}
Mydf$Date <- as.Date(with(Mydf,paste("1973",Month,Day,sep='-')), "%Y-%m-%d")
```

## Create line chart for ozone
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Date, y=Ozone, group=1)) + 
  geom_line(color="red", na.rm = TRUE) 
```

## Create line chart for temp
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Date, y=Temp, group=1)) + 
  geom_line(color="blue", na.rm = TRUE) 
```

## Create line chart for wind
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Date, y=Wind, group=1)) + 
  geom_line(color="green", na.rm = TRUE)
```

## Create line chart for solar.R
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Date, y=Solar.R, group=1)) + 
  geom_line(color="black", na.rm = TRUE)
```

## Combine 4 line charts
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf, aes(x=Date, group=1)) +
  geom_line(aes(y=Ozone, color="Ozone"), na.rm = TRUE) +
  geom_line(aes(y=Temp, color="Temp"), na.rm = TRUE) +
  geom_line(aes(y=Wind, color="Wind"), na.rm = TRUE) +
  geom_line(aes(y=Solar.R,color="Solar.R"), na.rm = TRUE) +
  scale_color_discrete(name="Lines") +
  ggtitle("Ozone, Temp, Wind and Solar.R vs Date Line Chart") +
  theme(plot.title=element_text(hjust=0.5)) +
  ylab("Values")
```

# Step 4: Look at all the data via a Heatmap
```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE}
library(reshape2)
mydf.m<- melt( Mydf, id.vars=c("Date"), names(Mydf)[1:4] )
mydf.m$scaled<- scale(mydf.m$value)
ggplot( data=mydf.m, aes(Date, variable)) + 
  geom_tile(aes(fill = scaled), na.rm = TRUE) + 
  scale_fill_gradient(low = "white", high = "steelblue")
```

# Step 5: Look at all the data via a scatter chart
```{r fig.width = 5, fig.height = 4}
ggplot(Mydf,aes(x=Wind, y=Temp)) + 
  geom_point(aes(color=Solar.R, size=Ozone), na.rm = TRUE) +
  ggtitle("Scatter Plot for Wind, Temp, Ozone and Solar.R") +
  theme(plot.title=element_text(hjust=0.5))
```

# Step 6: Final Analysis
• Do you see any patterns after exploring the data?  
Patterns:  
1. When the temperature goes up, ozone goes up as well.  
2. Temperature goes up when Solar Radiation is high.  
3. Ozone and temperature are high in September  

• What was the most useful visualization?  
Scatter plot is the most useful visualization. Because it can not only show the relationship between two variables but also give the distribution information.