---
title: "Bing-Je_WU_HW8"
author: "Bing-Je Wu"
date: "6/1/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Step 1. Load the data set
Download data set.
The first column shows the number of fawn in a given spring (fawn are baby Antelope).
The second column shows the population of adult antelope.
The third shows the annual precipitation that year.
Finally, the last column shows how bad the winter was during that year.
```{r , message=FALSE, warning=FALSE}
library(gdata)
mydf <- read.xls("mlr01.xls")
mydf
```

# Step 2. 
Inspect	the	data using the str() command to	make sure	that all of the	cases	have been	read in	(n=8 years of observations) and	that there are four variables.	
```{r , message=FALSE, warning=FALSE}
str(mydf)
antelopeDF <- data.frame(mydf$X1, mydf$X2, mydf$X3, as.factor(mydf$X4))
str(antelopeDF)
colnames(antelopeDF) <- c("baby","adult","precipitation","winter")
```

# Step 3. 
Create bivariate plots of	number of	baby fawns versus	adult	antelope population,	the	precipitation	that year, and the severity	of the winter.
Your code	should produce three separate	plots. Make	sure the Y-axis	and	X-axis are labeled. Keeping in	mind that	the	number of	fawns is the outcome (or dependent) variable,	which	axis should	it go	on in	your plots?
```{r , message=FALSE, warning=FALSE}
library(ggplot2)
 # Baby antelope vs. adult antelope
ggplot(antelopeDF, aes(x=adult, y=baby)) +geom_point() + xlab("Adult antelope") +
  ylab("Baby antelope") + ggtitle("Baby antelope vs. Adult antelope population") + 
  theme(plot.title = element_text(hjust = 0.5))
# Baby antelope vs. precipitation
ggplot(antelopeDF, aes(x=precipitation, y=baby)) +geom_point() + xlab("Precipitation") +
  ylab("Baby antelope") + ggtitle("Baby antelope vs. Precipitation") + 
  theme(plot.title = element_text(hjust = 0.5))
 # Baby antelope vs. Severity	of the winter
ggplot(antelopeDF, aes(x=winter, y=baby)) +geom_point() + 
  xlab("Severity of the winter") + ylab("Baby antelope") + 
  ggtitle("Baby antelope vs. Severity of the winter") + 
  theme(plot.title = element_text(hjust = 0.5))
```

# Step 4. 
Next,	create three regression	models of increasing complexity	using lm().
In the first model, predict the	number of	fawns	from the severity	of the winter.
In the second model, predict the number	of fawns from	two	variables	(one should	be the severity of	the	winter). In	the	third	model	predict	the	number of	fawns from the three other	variables. 
Which model works best? Which of the predictors are statistically significant in each model? 
If	you	wanted to	create the most	parsimonious model (i.e., the one that did the best job with	the	fewest predictors),	what would it	contain?

Fisrt model 
```{r , message=FALSE, warning=FALSE}
model_1 <- lm(formula = baby~winter,data = antelopeDF)
summary(model_1)
 # There is no predictor is statistically significant.
```

Second model
```{r , message=FALSE, warning=FALSE}
model_2 <- lm(formula = baby~winter + adult,data = antelopeDF)
summary(model_2)
 # The number of adult antelopes is statistically significant.
 # The adjust R-square 0.9379. It means 94% of y values can be explained by X variables.
 # This model works best among three models.
```

Third model
```{r , message=FALSE, warning=FALSE}
model_3 <- lm(formula = baby~winter + adult + precipitation, data = antelopeDF)
summary(model_3)
 # There is no predictor is statistically significant.

```

The most parsimonious model
```{r , message=FALSE, warning=FALSE}
model_4 <- lm(formula = baby~adult, data=antelopeDF)
summary(model_4)
 # Baay versus addult model will be the most parsimonious model. 
 # adult variable is statistically significant. R-squared is 0.8813. 
 # P-value is 0.0005 . The equation is also statistically significant. 
```
