---
title: "Bing-Je_Wu_HW2"
author: "Bing-Je Wu"
date: "4/14/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Copy the dataset from built-in mtcars dataset
```{r}
myCars <- mtcars
```

## Step 1. What is hp (horse power)
```{r finding the max hp from the dataset}
myCars[which.max(myCars$hp),]
```
1) What is the highest hp?
```{}
The highest hp is 335 (hourse power)
```
2) What car has highest hp?
```{}
Maserati has the highest house power among those vehicle
```


## Step 2. Explore mpg (miles per gallon)

```{r finding the max mpg from the dataset}
myCars[which.max(myCars$mpg),]
```
3) What is the highest mpg?
```{}
The highest mpg is 33.9 (miles per gallon)
```
4) What car has highest hp?
```{}
Toyota Corolla has the highest mpg among those vehicle
```
5) Create a sorted dataframe based on mpg
```{r}
head(myCars[order(myCars$mpg, decreasing = TRUE),], 10)
myCars_by_mpg <- myCars[order(myCars$mpg, decreasing = TRUE),]
```


## Step 3. Which car has the best combination of hp and mpg?
```{r filter the dataset}
myCars_by_mpg[myCars_by_mpg$mpg > mean(myCars_by_mpg$mpg),c(1,4)]
```
6) What logic did you use?
```{}
For me, mpg should be put on the first priorty for considering the best combination. 
So, I will calculate the mean of the mpg fisrt and filter out other data with lower mpg. 
Then, among the remaining data, find the maximum value of hp and get the best combination.
```
7) Which car?
```{}
Lotus Europa has the best combination of hp and mpg.
```


## Step 4. Which car has the "best" combination of hp and mpg where mpg and hp must be given equal weight?
```{r geting the quartile information}
summary(myCars)
```
```{}
The medians for mpg and hp are 19.2 and 123.0.
```
```{r}
myCars_by_mpg[,c(1,4)]
```
```{}
Thus, "Merc 280" is the car has the best combination fo hp and mpg with equal weight.
```
