---
title: "Bing-Je_Wu_HW4"
author: "Bing-Je Wu"
date: "4/25/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step1. Write a summarizing function to understand a distribution of a vector
```{r}
printVecInfo <- function(inputVector){
  # set mean, median, min ,max, and sd
  Vmean <- mean(inputVector)
  Vmedian <- median(inputVector)
  Vmin <- min(inputVector)
  Vmax <- max(inputVector)
  Vsd <- sd(inputVector)
  # set quantile on 0.05% and 0.95%
  Q005<- quantile(inputVector,c(0.05,0.95))[1]
  names(Q005)<-NULL
  Q005
  Q095<- quantile(inputVector,c(0.05,0.95))[2]
  names(Q095)<-NULL
  Q095
  # set Skewness
  library(moments)
  Vskewness <- skewness(inputVector)
  # print all values
  print(paste("mean:", Vmean))
  print(paste("median:", Vmedian))
  print(paste("min:", Vmin, "max:", Vmax))
  print(paste("sd:", Vsd))
  print(paste("quantile (0.05 ‐ 0.95):",Q005,"--",Q095))
  print(paste("Skewness:", Vskewness))
}
```


## Step 2. Creating samples in a Jar
*4. Create a variable 'jar' that has 50 red and 50 blue marbles*
```{r}
jar <- c(replicate(50,"red"), replicate(50, "blue"))
```
*5. Confirm it has 50 reds by summing the samples that are red*
```{r}
length(jar[jar == "red"])
```
*6. Sample 10 'marbles' from the jar. How many are red? What was the percentage of red marbles?*
```{r}
sample10<- sample(x = jar, size = 10, replace = TRUE)
length(sample10[sample10 == "red"])
length(sample10[sample10 == "red"])/length(sample10)
```
*7. Do the sampling 20 times, using 'replicate' command. This should generate a list of 20 numbers. Each number is the mean of how many reds there were in 10 samples. Use your printVecInfo to see information of the samples. Also generate a histogram of the samples.*
```{r}
listOf20<- 
  replicate(20,length(sample(jar,10,replace = TRUE)[
  sample(jar,10, replace = TRUE) == "red"])/length(
    sample(jar,10, replace = TRUE))
  )
printVecInfo(listOf20)
```
```{r listOf20 hist, fig.width = 5, fig.height = 4}
hist(listOf20)
```

*8. Repeat #7, but this time, sample the jar 100 times.You should get 20 numbers, this time each numebr represents the mean of how many reds were in the 100 samples. Use your printVecInfo to see information of the samples.Also generate a histogram of the samples.*
```{r}
listOfNew20 <- 
  replicate(20,length(sample(jar,100,replace = TRUE)[
  sample(jar,10, replace = TRUE) == "red"])/length(
    sample(jar,10, replace = TRUE))
)
printVecInfo(listOfNew20)
```
```{r listOfNew20 hist, fig.width = 5, fig.height = 4}
hist(listOfNew20)
```

*9. Repeat #8, but this time, replicate the sampling 100 times.You should get 20 numbers, this time each numebr represents the mean of how many reds were in the 100 samples. Use your printVecInfo to see information of the samples. Also generate a histogram of the samples.*
```{r}
listOfNew20with100times <- 
  replicate(100,length(sample(jar,100,replace = TRUE)[
    sample(jar,10, replace = TRUE) == "red"])/length(
      sample(jar,10, replace = TRUE))
  )
printVecInfo(listOfNew20with100times)
```
```{r listOfNew20with100times hist, fig.width = 5, fig.height = 4}
hist(listOfNew20with100times)
```


## Step 3. Explore the airquality dataset
*10. Store the 'airquality' dataset into a tempary variable*
```{r}
T_airquality <- airquality
```
*11. Clean the dataset (remove the NAs)*
```{r}
summary(T_airquality)
colSums(is.na(T_airquality))
 # Remove the rows with NA
T_airquality_clean <- na.omit(T_airquality)
  # Check NAs again
rownames(T_airquality_clean)<-NULL
any(is.na(T_airquality_clean))
colSums(is.na(T_airquality_clean))
```
*12. Explore Ozone, Wind and Temp by doing a ‘printVecInfo’ on each as well as generating a histogram for each*
```{r Ozone}
printVecInfo(T_airquality_clean$Ozone)

```
```{r Ozone hist, fig.width = 5, fig.height = 4}
hist(T_airquality_clean$Ozone)
```
```{r Wind}
printVecInfo(T_airquality_clean$Wind)

```
```{r Wind hist, fig.width = 5, fig.height = 4}
hist(T_airquality_clean$Wind)
```
```{r Temp}
printVecInfo(T_airquality_clean$Temp)

```
```{r Temp hist, fig.width = 5, fig.height = 4}
hist(T_airquality_clean$Temp)
```

