---
title: "Bing-Je_Wu_HW3"
author: "Bing-Je Wu"
date: "4/19/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE)
```

## #Step 1. Create a function (readStates) to read a CSV file into R
```{r store url}
urlRemote <- "https://www2.census.gov/"
path <- "programs-surveys/popest/tables/2010-2011/state/totals/"
fileName   <- "nst-est2011-01.csv"
urlToRead <- paste0(urlRemote, path, fileName)

```
```{r create readStates function}
readStates <- function(inputURL){
  library (RCurl)
  Temp <- getURL(inputURL)
  return(read.csv(text=Temp))
}
```
```{r message=FALSE, warning=FALSE}
mytable <- readStates(urlToRead)
```


## #Step 2. Clean the dataframe
Remove empty columns:
```{r Remove empty columns}
mytable <- mytable[,1:5]
```
Remove top 8 rows:
```{r Remove top 8 rows}
mytable <- mytable[-1:-8,]
rownames(mytable)<-NULL
```
Remove bottom 6 rows:
```{r Remove bottom 6 rows}
mytable <- mytable[-52:-58,]
```
Rename column, remove the old column, and normoalize it:
```{r Rename column, remove the old column, and normoalize it}
mytable$stateName <-mytable[,1]
mytable <- mytable[,-1]
mytable$stateName <- gsub("\\.","",mytable$stateName)
```
Normalize X,X.1,X.2,X.3 variables:
```{r Normalize X,X.1,X.2,X.3 variables}
mytable$base2010 <- gsub("\\,","",mytable$X)
mytable$base2010 <- as.numeric(mytable$base2010)

mytable$base2011 <- gsub("\\,","",mytable$X.1)
mytable$base2011 <- as.numeric(mytable$base2011)

mytable$Jul2010 <- gsub("\\,","",mytable$X.2)
mytable$Jul2010 <- as.numeric(mytable$Jul2010)

mytable$Jul2011 <- gsub("\\,","",mytable$X.3)
mytable$Jul2011 <- as.numeric(mytable$Jul2011)
```
Remove old X, X.1, X.2, X.3 columns:
```{r Remove old X, X.1, X.2, X.3 columns}
mytable <- mytable[,-1:-4]
```
Analyze the dataset:
```{r}
summary(mytable)
str(mytable)
```

## #Step 3. Store and explore the dataset
Store the dataset as dfStates:
```{r store the dataset as dfStates}
dfStates <- mytable
```
Calculate the mean for the July2011 data:
```{r calculate the mean for the July2011 data}
mean(dfStates$Jul2011)
```

## #Step 4: Find the state with the Highest Population

```{r }
dfStates[which.max(dfStates$Jul2011),]
```
Based on the July2011 data, California is the state that has the highest population.
```{r}
dfStates[order(dfStates$Jul2011, decreasing = FALSE),]
```

## #Step 5: Explore the distribution of the states
Write a function:
```{r function}
below_percentage <- function(inputvector, inputnumber){
  Total_number <- length(inputvector)
  Number_below <- length(inputvector[inputvector<inputnumber])
  return(Number_below/Total_number)
}
```
Test the function:
```{r Test the function}
A = c(1,2,3,4,5)
a = 2
below_percentage(A,a)
```
Test vector 'dfStates$Jul2011Num' and the mean of dfStates$Jul2011Num':
```{r }
below_percentage(dfStates$Jul2011, mean(dfStates$Jul2011))
```

